---
title: "p8105_hw2_zx2424"
output: github_document
author: "Zhaoqianyu Xiong"
---
```{r}
library(tidyverse)
```

##Problem 1
The dataset after selection contains variables: line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Now, I have imported data, clean variables names, select required data, and convert entry variable from character to logical variable. The resulting dataset is 1868*19. These data are not tidy.
```{r}
subway_data = read_csv("./data/NYC_Transit_Subway.csv",
                       col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

From results of the code below, there are 465 distinct stations.
```{r}
subway_data %>%
  select(line, station_name) %>%
  distinct
```

From results of the code below, 84 station are ADA compliant.
```{r}
subway_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Variable "entry" is now logic, so we only need to take the mean. And the mean will be the proportion. The proportion of entrances without vending is 0.377.
```{r}
subway_data %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

There are 60 distinct stations serve the A train.
```{r}
subway_data %>%
pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>%
  filter(route == "A") %>%
  select(line, station_name) %>%
  distinct
```

There are 17 stations of the ones serve the A train ADA compliant.
```{r}
subway_data %>%  
pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

##Problem 2 
```{r}
library(readxl)
excel_sheets("./data/Trash Wheel Collection Data.xlsx")
```

```{r}
Mr_Trash_Wheel = read_excel("./data/Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N550" ) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(Wheel_type = "Mr. Trash Wheel")  %>%
  mutate(dumpster = as.character((dumpster))) %>%
  mutate(year = as.integer(year))
```

```{r}
Professor_Trash_Wheel = read_excel("./data/Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M97") %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(dumpster = as.character(dumpster)) %>%
  mutate(Wheel_type = "Professor Trash Wheel") %>%
  mutate(sports_balls = 0)
```

```{r}
Trash_Wheel = 
  bind_rows(Mr_Trash_Wheel, Professor_Trash_Wheel) %>%
  janitor::clean_names()
```

The resulting dataset after combination shows the total number of dumpsters consumed is 568. Many different kinds of trash are collected, including plastic bottles, polystyrene, cigarette butts, glass bottles, chip bags and sports balls. And the total weight of trash collected by two different trash wheels is `r sum(pull(Trash_Wheel, weight_tons))` tons. 
The total weight of trash collected by Professor Trash Wheel is `r sum(pull(Professor_Trash_Wheel, weight_tons))` tons. The total number of sports balls collected by Mr. Trash Wheel is `r sum(pull(Mr_Trash_Wheel, sports_balls))`.

## Problem 3
Clean the data in pols-months.csv.
```{r}
pols_month = read_csv("./data/pols-month.csv") %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(year = as.integer(year)) %>%
  mutate(month = as.integer(month)) %>%
  mutate(month = month.abb[month]) %>%
  mutate(president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-prez_dem) %>%
  select(-prez_gop) %>%
  select(-day)
```

Clean the data in snp.csv.
```{r}
snp = read_csv("./data/snp.csv") %>%
    separate(col = date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(year = as.integer(year)) %>%
  mutate(month = as.integer(month)) %>%
  mutate(year = ifelse(year < 50, 2000 + year, 1900 + year)) %>%
  arrange(year, month) %>%
  mutate(month = month.abb[month]) %>%
  select(year, month, everything()) %>%
  select(-day)
```

Tidy the unemployment data.
```{r}
unemployment = read_csv("./data/unemployment.csv") %>%
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "percentage_of_unemployment") %>%
    rename(year = Year) %>%
    mutate(year = as.integer(year))
    
```

Join the datasets.
```{r}
FiveThirtyEight = 
  left_join(pols_month, snp, by = c("year", "month")) %>%
  left_join(unemployment, by = c("year", "month"))

FiveThirtyEight
```

The pols_month dataset is 822*9 after cleaning, containing the information of which party the president belonged to, the number of republican or democratic governers, senators and representatives from 1947 to 2015.
The snp dataset is 787*3 after cleaning, containing the information about the closing values of the S&P stock index from 1950 to 2015.
The unemployment dataset is 816*3 after cleaning, containing the information concerning the percentage of unemployment from 1948 to 2015.
The resulting dataset "FiveThirtyEight" is 822*11, containing the all information of three datasets mentioned above. And its range of years is 1947 to 2015.

